---
title: "GSS challange"
author: "Emann Malik"
date: "March 27, 2021"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(knitr)
GSS <-read.csv(file="GSScsv.csv",header=TRUE)
```
  
```{r echo=FALSE}
GSS$RACE <- as.factor(GSS$RACE)
GSS$BIGBANG <- as.factor(GSS$BIGBANG)
```
  
# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

Is belief in the bigbang theory (BIGBANG) independent of race (RACE)?
If not, then whose view differs from whom??
  
In this report, we will be trying to show if there is a relationship at all between a person's race and whether or not they believe in the Big Bang theory. The data to be used comes from the GSS data set. From this data set, we will be focusing on the factor variables RACE (with the three levels BLACK, WHITE, and OTHER) and BIGBANG (with the two levels False and True). We will be attempting to answer this research question: Is the belief in the Big Bang theory independent of race?
  
To help answer this research question, we will take two hypotheses into account; the null, which states that the belief in the Big Bang theory is independent of race, and the alternative, which states that the belief in the Big Bang theory is not independent of race.
  
# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->
  
We will be using a variable analysis for this report. The graphical method of a bar plot, numerical method of percent tables, and the inferential tools of the Chi-squared test and Fisher's exact test (both normal and 2x2) will be utilized to conduct this variable analysis. 
  
# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->
  
```{r}
dd2 <- GSS %>% group_by(RACE, BIGBANG) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=RACE,y=count,fill=BIGBANG))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=RACE,y=prcnt*100,fill=BIGBANG)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
  
The bar plots above show anything but uniform beliefs in the Big Bang theory across different races. Blacks have the most extreme differences in opinion, with a whopping 81% believing that the Big Bang theory is false and only about 19% believing it is true. Whites have a completely different variation of opinion, with half believing in the theory and half not believing in it. In the middle, other races beside black and white also have a unique variation of opinion in comparison to the other two races, with about 56% believing the Big Bang theory is true and 44% believing it is not true. We are starting to see that the belief in this theory differs among multiple races.
  
### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
  
```{r}
table2 <- xtabs(~RACE + BIGBANG, data=GSS)
rowPerc(table2)
colPerc(table2)
```
  
These percent tables show quite a significant difference between the three categories of races in this data set. Once again, we see that the numbers vary drastically among blacks, whites, and other races: most blacks (approximately 81%) do not believe the Big Bang theory is true, whites are split between believing and not believing in the theory (50% and 50%), and the "other" category contains more people who believe in the Big Bang theory (approximately 56%) than those who don't (approximately 44%).
  
## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->
  
```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```
  
The hypotheses for this analysis are as follows:  
null: The belief in the Big Bang theory is independent of race.  
alternative: The belief in the Big Bang theory is not independent of race.  
  
The Chi-squared test above displays a p-value of 0.01098 while Fisher's exact test displays a p-value of 0.00905. Both of these values, which represent the chance that we will get the same results just received if the null hypothesis is true, are small numbers. Because of this, we reject the null in favor of the alternative: the belief in the Big Bang theory is not independent of race. 
  
## A function to run multiple fisher tests  
### Upload the fisherXY.R function then source it

```{r}
fisherXY = function(x, y, confL = 0, printAll=FALSE){
  # Inputs: 
  #   x: factor variable, 2 levels
  #   y: factor variable, n levels
  # confL: confidence level, default = 0.95
  #        print only those tables, results if p-value < (1-confL)
  # printAll: overide the 'print only if p-value < (1-confL), and
  #           print all tables and fisher test results
  #
  # Constraint:
  #   x and y must be same length
  #
  # Outputs:
  #   All 2x2 Fisher exact tests
  
  if(confL<0 || confL> 1){
    print('error: confL must be between 0 and 1')
    return(999)
  }
  
  if(!is.factor(x)) x=as.factor(x)
  if(!is.factor(y)) y=as.factor(y)
  
  fisherTable = as.matrix(table(x, y))
  cat('\ncontingency table:\n')
  print(fisherTable)
  cat('\n***All 2X2 tests****\n\n')

  # all 2x2 tests
  levelsY = levels(y)
  levelsX = levels(x)
  
  m = length(levelsX)
  n = length(levelsY)
  
  confLevel = 0.95
  if (confL !=0 ) confLevel=confL
  
  for (i in 1:(m-1))
       for(j in ((i+1):m))
          for (k in 1:(n-1))
             for (l in ((k+1):n)) {
                
                 fisherIndexX = c(i,j)
                 fisherIndexY = c(k,l)
                 
                 fTempTable = as.matrix(fisherTable[fisherIndexX,
                                        fisherIndexY])
                 
                 fisherTest=fisher.test(fTempTable, conf.int = TRUE, 
                              conf.level = confLevel)
                 
                 if((fisherTest$p.value < (1-confLevel) ) || printAll){
                   
                       print(c(levelsX[i], levelsX[j]))
                       print(c(levelsY[k], levelsY[l]))
                       print(fTempTable)
                       print(fisherTest)
                       cat('\n')
                 }
                 
        }
}

```
```{r}
fisherXY(GSS$RACE,GSS$BIGBANG)
``` 
  
The multiple Fisher tests above begin by comparing blacks and "other." The results of the test displays a p-value of 0.02252 and a confidence interval of 1.144365 to 25.429759. This low p-value and confidence interval with no 1 in it signify that blacks and "other" are significantly different in their views on the Big Bang theory. 
  
The next test, between black and white views, shows that they too are significantly different (even more than blacks and "other" are different) with a low p-value of 0.004628 and confidence interval of 1.418155 to 14.982411, with no 1. 
  
Once again, we have compelling evidence that the null hypothesis is to be rejected in favor of the alternative (The belief in the Big Bang theory is not independent of race.), since two pairs of compared races are significantly different in terms of views on the Big Bang theory.
  
# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
